---
title: "R Objects"
subtitle: "STAT486/586"
author: "Xiongtao Dai"
output:
  xaringan::moon_reader:
    css: [myslides.css, myslides-fonts.css]
    lib_dir: libs
    nature:
      beforeInit: "macros.js"
      highlightLines: false
      countIncrementalSlides: false
---

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```


## Outline

- The type system

---
class: middle

# John Chamber's 1st Principle: Everything that exists in R is an object

---

## R types

- Literally, everything in R is an object. 

- The *type* of an object refers to the internal storage mode. Shown by `typeof()`

- The *class* of an object determines the how other functions interpret the object. Used by S3 method dispatch. Shown by `class()`

- There are only a few R data types. Most common are vectors
    - double
    - integer
    - character
    - logical
    - list

---

## Attributes

- An R object is supplemented with attributes, e.g., `names` and `class`
    ```{r}
    x <- c(5, 10)
    names(x) <- c("cat", "dog")
    class(x) <- "PetCounts"
    attributes(x)
    ```

- Attributes are usually not preserved by R functions
    ```{r}
    attributes(c(x))
    ```

---

## Building R objects

- More complex objects can be built from simple ones by adding attributes
    ```{r}
    a <- 1:26
    str(a)
    attr(a, "levels") <- letters
    attr(a, "class") <- "factor"
    str(a)
    ```


- S3 objects are built this way. E.g., results from `lm` models, data frames

---
class: inverse

## Your Turn

- Define a matrix `x <- matrix(1:6, 2, 3)`. Inspect the attributes of `x` by calling `attribute`

- Add attributes to a vector `y <- 1:6` until it becomes the same as `x` (without using the `matrix` function). Print out the resulting object

---
## More objects

- More objects we will encounter in this course

    - Functions
    
    - Environments. E.g. .GlobalEnv
    
    - S4 object. E.g., results returned by `stats4::mle`
    
    - Language. E.g., `y ~ x`
	

