---
title: "Making code more reliable"
subtitle: "PH 290"
author: "Xiongtao Dai"
output:
  xaringan::moon_reader:
    css: [myslides.css, myslides-fonts.css]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
      highlightLines: false
      countIncrementalSlides: false
---

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```


## Outline

- Writing and documenting functions
    
- Unit testing

- Exceptions and handling

- Coding style

---

## How to write good functions

- Think clearly about the purpose/scope (what it does and what it does not)

- Think about the input/output, i.e., Application Programming Interface, (API). 

    - APIs are hard to change without breaking compatibility, so anticipate future demands

-  Be intuitive

-  Be consistent

- Read other's code

- Write (a lot of) code. Maintain your software

---

## Documenting functions

- Documentation helps the future you, and anyone who will potentially use your function

- The documentation of a function should describe

    1. The purpose of the function
    
    1. Inputs. This should include argument names and the expected formats of the inputs
    
    1. Outputs and their format
    
    1. Better still, some details of the implementation/algorithm and list some example usages
    

---
class:big, middle

## Testing functions

---

## Unit testing

- *Unit testing* is the procedure of testing a small unit of code. A unit is often a function

- Correctness is the #1 priority. Having unit tests improves software reliability

- Functions and tests are like duos. They improve each other

- Prevents regression, namely old functionality stops to work after new ones are introduced

- Writing a unit test is not difficult at all. After you finish a freshly written function, you must have written some code to see whether it works or not. *Do not discard the code*, but rather make it a formal unit test

---

## The `testthat` package

- `testthat` is a package for unit testing your code. Its goal is to make testing rigorous and fun 

    ```{r}
    library(testthat)
    ```

- The main workhorse is the expectation/assertion function

    ```
    expect_equal(object, expected, ..., tolerance)
    ```

---

## `expect_equal`

```
expect_equal(object, expected, ..., tolerance)
```

- `expect_equal` tests whether the output of your code matches your expectation

- If matches, nothing happens (yeah!)

- If not, you have an error, so you must go back and check your code or the test


---

- A very basic example:
    ```{r}
    x <- 1:3
    expect_equal(mean(x), 2)
    ```

- By default, differences below a small threhold will be regarded OK. This is useful when comparing floating point numbers
    ```{r, error=TRUE}
    y <- seq(-100, 100, by=0.1)
    expect_equal(mean(y), 0) # while `mean(y) == 0` is FALSE

    ```

- Set the tolerance threshold using `tolerance=`
    ```{r, error=TRUE}
    expect_equal(mean(y), 0, tolerance = 1e-16)
    ```

---

- Group related unit tests using `test_that(desc, code)`

- `test_that()` also creates an isolated context to run the test

```{r, eval=FALSE}
myMean <- function(x, na.rm=FALSE) {
  if (na.rm) {
    x <- x[!is.na(x)]
  }
  
  sum(x) / length(x)
}

test_that("myMean works", {
  x <- 1:3
  y <- seq(-100, 100, by=0.1)
  z <- c(y, NA)
  expect_equal(myMean(x), 2)
  expect_equal(myMean(y), 0)
  expect_equal(myMean(z), NA) # NA has types!
  expect_equal(myMean(z, na.rm=TRUE), 0)
})
```
---

## A suite of `expect_*` functions

- `expect_identical()`

- `expect_true()`, `expect_false()`

- `expect_lt()`, `expect_lte()`, `expect_gt()`, `expect_gte()`

- `expect_type()`, `expect_s3_class()`, `expect_s4_class()`

- `expect_error()`, `expect_warning()`, `expect_message()`

- `expect_snapshot()`

and [more](https://testthat.r-lib.org/reference/index.html)

 
---
## Writing good tests

A good unit test checks whether a *specific aspect* of your code meets your expectation. For example,

- the output has the expected type, length, or dimension

- the output equals what you expect in a simple/known case

- the output behaves correctly in corner cases, such as when $n=1$

- an optional argument works


---

class: inverse

## Your turn

- Add a test for `myMean` to check that the function works if `x` has length 1

---

## Assertions

- `expect_equal()` is an example of *assertion* functions, which does not do anything if expected conditions are met or otherwise gives an error

- `stopifnot(condition)` is the buildin version that gives an error unless `condition` is `TRUE`

- Assertions are also useful when writing a function or an analysis for capturing common mistakes

- Making your function fail faster helps pinpointing the mistake
    ```{r}
    GenSample <- function(n) {
      stopifnot(n >= 1)
      # ...generate sample
    }
    ```
    ```{r, eval=FALSE}
    # not real code
    dat <- CleanData(dat)
    stopifnot(all(!is.na(dat$Sex)))
    # ...more analysis
    ```

---

## Conditions

<!-- - You can raise errors, warnings, and messages using `stop()`, `warning()`, and `message()` -->

- An error should be raised if the computation cannot continue
    ```{r, eval=FALSE}
    stop("Dimensions mismatch")
    ```

- A warning should be used if the computation can continue but the results may not be reliable
    ```{r, eval=FALSE}
    warning("The sample size `n` is too small")
    ```


- A message is for informing the user something was done on their behalf
    ```{r, eval=FALSE}
    message("Creating `figures/` directory")
    ```

---

## Ignoring conditions

```{r}
a <- function(x) {
  try(someCodeThatErr)
    x
} # Shows the error message, but continues to run
a("Oh no")

suppressWarnings(warning("global warming"))
suppressMessages(message("tic toc tic toc"))
```

---

## Handling conditions
- Handling error
    ```{r}
    tryCatch({
      res <- runAnAlgorithmThatMayErr
      print("algorithm finishes")
    }, error=function(x) print("algorithm cannot finish"))
    ```

- Handling warnings/messages
    ```{r}
    withCallingHandlers({
      warning("algorithm does not converge")
      NA
    }, warning=function(x) print("algorithm not converged"))
    ```

---

## Coding style

- Follow a good coding style, e.g., the [tidyverse style guide](https://style.tidyverse.org/)
    - Prioritize readability
    - Consistency
    - Good naming convention
    - Nice spacing/indentation
    - Use standard/conventional usage of functions

- Practice defensive coding
    - Program clearly and rigorously
        - Be clear about the argument/variable/function you are referring to
        - If unsure about the precedence, clarify using `()`
        - Use `drop=FALSE` when dealing with matrices and data frames: `aMatrix[, 1, drop=FALSE]`
    - Program parsimoniously
        - Don't load packages if not necessary
    - Use assertions
    - Use conditions
