---
title: "Making code more reliable"
subtitle: "STAT486/586"
author: "Xiongtao Dai"
output:
  xaringan::moon_reader:
    css: [myslides.css, myslides-fonts.css]
    lib_dir: libs
    nature:
      beforeInit: "macros.js"
      highlightLines: false
      countIncrementalSlides: false
---

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```


## Outline

- Writing and documenting functions
    
- Unit testing

- Exceptions and handling

- Coding style

---

## How to write good functions

- Think clearly about the purpose/scope (what it does and what it does not)

- Think about the input/output, i.e., Application Programming Interface, (API). 

    - APIs are hard to change without breaking compatibility, so anticipate future demands

-  Be intuitive

-  Be consistent

- Read other's code

- Write (a lot of) code. Maintain your software

---

## Documenting functions

- Documentation helps the future you, and anyone who will potentially use your function

- The documentation of a function should describe

    1. The purpose of the function
    
    1. Inputs. This should include argument names and the expected formats of the inputs
    
    1. Outputs and their format
    
    1. Better still, some details of the implementation/algorithm and list some example usages
    

---
class:big, middle

## Testing functions

---

## Unit testing

- *Unit testing* is the procedure of testing a small unit of code. A unit is often a function

- Correctness is the #1 priority. Having unit tests improves software reliability

- Functions and tests are like duos. They improve each other

- Writing a unit test is not difficult at all. After you finish a freshly written function, you must have written some code to see whether it works or not. *Do not discard the code*, but rather make it a formal unit test

---

## The `testthat` package

- `testthat` is a package for unit testing your code. Its goal is to make testing rigorous and fun 

    ```{r}
    library(testthat)
    ```

- The main workhorse is the expectation/assertion function

    ```
    expect_equal(object, expected, ..., tolerance)
    ```

---

## `expect_equal`

```
expect_equal(object, expected, ..., tolerance)
```

- `expect_equal` tests whether the output of your code matches your expectation

- If matches, nothing happens (yeah!)

- If not, you have an error, so you must go back and check your code or the test


---

- A very basic example:
    ```{r}
    x <- 1:3
    expect_equal(mean(x), 2)
    ```

- By default, differences below a small threhold will be regarded OK. This is useful when comparing floating point numbers
    ```{r, error=TRUE}
    y <- seq(-100, 100, by=0.1)
    expect_equal(mean(y), 0) # while `mean(y) == 0` is FALSE

    ```

- Set the tolerance threshold using `tolerance=`
    ```{r, error=TRUE}
    expect_equal(mean(y), 0, tolerance = 1e-16)
    ```

---

- Finally, group related unit tests using `test_that(desc, code)`:

```{r}
myMean <- function(x, na.rm=FALSE) {
  if (na.rm) {
    x <- x[!is.na(x)]
  }
  
  sum(x) / length(x)
}

test_that("myMean works", {
  x <- 1:3
  y <- seq(-100, 100, by=0.1)
  z <- c(y, NA)
  expect_equal(myMean(x), 2)
  expect_equal(myMean(y), 0)
  expect_equal(myMean(z), NA_real_) # NA has types!
  expect_equal(myMean(z, na.rm=TRUE), 0)
})
```
---

## A suite of `expect_*` functions

- `expect_identical()`

- `expect_true()`, `expect_false()`

- `expect_lt()`, `expect_lte()`, `expect_gt()`, `expect_gte()`

- `expect_type()`, `expect_s3_class()`, `expect_s4_class()`

- `expect_error()`, `expect_warning()`, `expect_message()`

- `expect_snapshot()`

and [more](https://testthat.r-lib.org/reference/index.html)

 
---
## Writing good tests

A good unit test checks whether a *specific aspect* of your code meets your expectation. For example,

- the output has the expected type, length, or dimension

- the output equals what you expect in a simple/known case

- the output behaves correctly in corner cases, such as when $n=1$

- an optional argument works


---
class: inverse

## Your turn

- Add a test for `myMean` to check that the function works if `x` has length 1
