---
title: "Scoping Rules"
subtitle: "PH 290"
author: "Xiongtao Dai"
output:
  xaringan::moon_reader:
    css: [myslides.css, myslides-fonts.css]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
      highlightLines: false
      countIncrementalSlides: false
---

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

class: big, middle

## How does R find global variables?

---

## How does R find global variables?

Use your intuition to predict what the following functions output:

1. 
    ```{r, eval=FALSE}
    x <- 1
    f1 <- function() {
      x <- 2
      print(x)
    }
    f1()
    ```

2. 
    ```{r, eval=FALSE}
    times <- 1
    f3 <- function() rep(0, times)
    g <- function() {
      times <- 2
      f3()
    }
    g()
    ```
---

3\. 

```{r, eval=FALSE}
f2 <- function() {
  if (!exists("a")) {
    a <- 1
  } else {
    a <- a + 1
  }
  a
}
f2()
f2()
```
    
4\. 

```{r, eval=FALSE}
x <- 1
f_4 <- function() x
x <- 2
f_4()
```

--

At the end of the class, you will understand why the functions behave as such (and thus you can program better)

---
class: big, middle

## Lexical scoping

---

## Lexical scoping

- *Scoping* is the act of finding values associated with a variable name. It is about *where* and *when* does R find global variables

- R follows the *lexical scoping* rule (aka static scoping)
    - Most programming languages use lexical scoping, like Python and C++

- *Where* R looks up values of variable names depends on how a function is *defined*, not how it is *called*
    - This makes R code easier to reason with 
    - Very useful when used in conjunction with encapsulation

- This is as opposed to *dynamic scoping*, where the value of a global variable depends on how a function is called

---

## Four rules for lexical scoping

- There are four principles for lexical scoping:

    1. Where to lookup

    1. Functions are variables

    1. Fresh starts

    1. When to lookup

---


1. *Where to lookup* (a): Local variables defined inside a function are the most relevant and mask variables defined outside a function
    ```{r, eval=TRUE}
    x <- 1
    f1 <- function() {
      x <- 2
      print(x)
    }
    f1()
    ```

---

1. *Where to lookup* (b): If a variable isn’t defined inside a function, R looks one level up where the function is *defined*
    ```{r, eval=TRUE}
    x <- 1
    f1_1 <- function() {
      print(x)
    }
    f1_1()
    ```
    
    ... *not* where the function is *called*
    ```{r, eval=TRUE}
    times <- 1
    f_pkg <- function() rep(0, times)

    g_usr <- function() {
      times <- 2
      f_pkg()
    }
    g_usr()
    ```
---

1. *Where to lookup* (c): The lookup rule applies the same if a function is defined inside another function

```{r}
x <- 0
f1_2 <- function() {
  x <- 1
  g <- function() {
    y <- 2
    c(x, y)
  }
  g()
}
f1_2()
```

- By "where", I mean which environment should the function look up variables. More next time

---

2\. *Functions are ordinary variables*, so the scoping rules apply when looking up both functions and variables

```{r}
f1_pkg <- function() print(1)
f2_pkg <- function() f1_pkg()

g2_usr <- function() {
  f1_pkg <- function() print(2)
  f2_pkg()
}
g2_usr()
```

---

3\. *Fresh starts*: The execution of a function is not affected by the previous executions
```{r}
f3 <- function() {
  if (!exists("a")) {
    a <- 1
  } else {
    a <- a + 1
  }
  a
}
f3()
f3()
```

---

4\. *When to lookup*. R looks up the values when the function is *called*, not when the function is *defined*. This is referred to as *Dynamic lookup*

```{r}
x <- 1
f_4 <- function() x
x <- 2
f_4()
```

- Rule 1 determines *where*, and Rule 4 determines *when* to look for values


---
class: big, middle

## Implications of lexical scoping

---

## Implications of lexical scoping

1. Encapsulation: Prevent users to modify the behavior of code through defining their versions of global variables/functions

- Your variable names won't clash with local variable names in another function, so the following code will work
    ```{r}
    myMap <- function(.l, .f, ...) {
        len <- length(.l)
        for (i in seq_len(len)) {
            print( .f(.l[[i]], ...) )
        }
    }
    len <- 5
    myMap(list(3, 2), function(x) rep(x, len))
    ```

- But be wary about the (input) argument names clashes

---

- Function factory works for the same reason
    ```{r}
    power1 <- function(exp) {
      function(x) {
        x ^ exp
      }
    }
    square <- power1(2)
    exp <- 3
    square(2)
    ```

---

2\. Be able to let function depend on system statuses at either *definition time* or *run time*

- A function that returns the time when it is defined:
    ```{r}
    getDefTime <- function() {
        time <- Sys.time()
        function() time
    }
    definitionTime <- getDefTime()
    definitionTime()
    ```

- A function that returns the time when it is called:
    ```{r}
    runTime <- Sys.time
    runTime()
    ```

- The same principle applies to tell package versions are available (if you compile an R package which reproduces an analysis), hardware resources, etc

---

## When is dynamic scoping useful?

- Relatively few languages use dynamic scoping, e.g., LaTeX and Bash

- Dynamic scoping allows users to modify the "state" and thus the behavior of software easily

---

## References

- Chapters 6, [Advanced R](https://adv-r.hadley.nz/index.html) by Hadley Wickham
