---
title: "Lazy Evaluation"
subtitle: "PH 290"
author: "Xiongtao Dai"
output:
  xaringan::moon_reader:
    css: [myslides.css, myslides-fonts.css]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
      highlightLines: false
      countIncrementalSlides: false
---

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

## Lazy evaluation 

- In R, function arguments are lazily evaluated, namely they are evaluated only if they are accessed.

- *Lazy evaluation* is as opposed to *eager evaluation*, the latter of which means an expression is evaluated as soon as it is written

- Benefits of lazy evalution:

    - Save computation

    - Allows for flexible programming

---

## Lazy evaluation principles
There are three principles:

1. Function arguments are held as an unevaluated *expression*, a.k.a., a *promise*. They are evaluated only if accessed

- You don't have an error because the argument `x` is not used in `h()`
    ```{r, error=TRUE}
    h <- function(x) {
      10
    }
    h(stop("an error here"))
    ```

---

2\. The evaluation is done only once when the value is accessed, and then it is cached (remembered). So `double(20)` in the following example is evaluated only once

```{r}
double <- function(x) { 
  message("Calculating...")
  x * 2
}

h2 <- function(x) {
  c(x, x)
}

h2(double(20))
```

---

3\. The evaluation is evaluated in the environment *where the expression is defined*

```{r}
h3 <- function(x) {
  y <- 100
  x
}

y <- 1
h3(print(y))
```

- This means users cannot inject code into a function through the input


---

## Some applications of lazy evaluation

Lazy evaluation is useful for the following tasks:

1. Specifying default arguments for a function

1. Non-standard evaluation

---

## Default arguments

- The default argument of a function can depend on another argument. E.g., see `?seq`

    A simplified e.g.
    ```{r}
    h4 <- function(x = 2 * y, y) {
      c(x, y)
    }
    h4(y=1)
    ```
    
    We do not have an error because `x = 2 * y` is no evaluated at the beginning of the function, but only evaluated at `c(x, y)`

---

## Default arguments

- However, the default argument is lazily evaluated *within* the function body, not at where the function is defined

```{r}
f <- function(x=g(y)) {
    y=1; print(x)
}
y <- 2
g <- identity
```
```{r}
f()
f(g(y))
```

- Avoid complex usage of the default argument, because it is hard for the users to reason with the output without looking at the body of the function

---

## Missing arguments

- We can handle missing arguments without default values

```{r}
h5 <- function(x) {
  if (missing(x)) {
    x <- "my default"
  }
  x
}
h5()
```

---

## Non-standard evaluation

*Standard evaluation* rules interpret arguments by the values

*Non-standard evaluation* (NSE) evaluates according to the expression but not the value of the expression


<!-- - `library` is NSE in its first argument -->
<!--     ```{r, error=TRUE} -->
<!--     library(MASS) # OK -->
<!--     x <- "MASS" -->
<!--     library(x) --> 
<!--     ``` -->


- E.g. of NSE: `plot` is NSE in the `x` and `y` arguments
    ```{r, fig.height=4, fig.width=5}
    myX <- 1:10
    plot(myX, sin(myX)) # How does plot() set the label names?
    ```

---

## References

- Chapters 6, [Advanced R](https://adv-r.hadley.nz/index.html) by Hadley Wickham
