---
title: "S4 and R6"
subtitle: "PH 290"
author: "Xiongtao Dai"
output:
  xaringan::moon_reader:
    css: [myslides.css, myslides-fonts.css]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
      highlightLines: false
      countIncrementalSlides: false
---

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

## Outline

- The S4 system

- The R6 system

---

class: big, middle

## S4

---

## S4

- S4 is a formal class system in R

- Has validator, multiple inheritance, multiple dispatch, etc for more rigorous and featureful OOP

- All related functions are in `library(methods)` (loaded by default)

---

- S4 has a formal definition of class using `setClass()` 
    ```{r}
    setClass("Patient", 
             slots = c(ID = "integer", pname = "character"), 
             prototype = list(ID = integer(), pname=character()))
    ```

- `showClass()` prints the information
    ```{r}
    showClass("Patient")
    ```


---

## Constructor

- `new()` is a low-level constructor for the class. Should be used by only the developer
    ```{r}
    new("Patient", ID=312L, pname="Max")
    ```

- We should make available a high-level constructor for the users
    ```{r}
    NewPatient3 <- function(ID, pname, ..., Class="Patient") {
        ID <- as.integer(ID)
        do.call(new, list(Class, ID=ID, pname=pname, ...))
    }
    p1 <- NewPatient3(1, "Max")
    ```

---

## Slots

- *Slots* of an S4 object (S3: fields) are accessed using `@` (S3: `$`) and `slot()` (S3: `[[`)
    ```{r}
    p1@pname
    slot(p1, "ID")
    ```

- Slot names are given by `slotNames()` (S3: `names`)


---

## Validator

- Validator is created using `setValidity()`. Returns `TRUE` if the object is valid, or a character vector for error messages. 
    ```{r, results="hide"}
    setValidity("Patient", function(object) {
        msg <- character()
        if (object@ID < 0L) {
            msg <- c(msg, "ID should be positive integer")
        } 
        if (nchar(object@pname) > 20) {
            msg <- c(msg, "pname should be shorter than 20")
        }

        if (length(msg) > 0) {
            return(msg)
        } else {
            return(TRUE)
        }
    })
    ```

- It will be run every time a new object is created via `new()`. Can also be invoked explicitely using `validObject()`

---

## Generic and methods

- `show()` is the S4 analogue of `print()` in S3

- `setMethod(generic, class, method)` sets the method
    ```{r}
    setMethod("show", 
              signature="Patient", 
              function(object) 
                  writeLines(sprintf("Patient %d is %s", 
                      object@ID, 
                      object@pname)))
    p1
    ```

- The class signiture can be `"ANY"` for a default method, or `"MISSING"` for a missing input

---

## Inheritance

- `setClass(..., contains=superclasses)` defines a subclass extending the given superclasses
    ```{r}
    setClass("In-patient", 
             contains = "Patient",
             slots = c(DOA = "Date"),
             prototype = list(DOA = as.Date(NA)))
    p3 <- NewPatient3(3, "Mary", 
                      DOA=as.Date("2022-01-02"), 
                      Class="In-patient")
    ```

- `contains="VIRTUAL"` creates a virtual class. This class is only for being inherited upon. No instance from a virtual class may be created

---

- `is(object)` reports the class, `is(object, class)` tests class membership
```{r}
is(p1)
is(p1, "In-patient")
```

---

## Coercion

- S4 has a powerful class coercion system
```{r}
p4 <- as(p1, "In-patient")
p4
```

- `setAs()` specifies a new coercion method
```{r}
setAs("Patient", "character", function(from) from@pname)
as(p4, "character")
```

---

## New generic

- `setGeneric(generic, definition)` creates a new generic

- `standardGeneric(generic)` starts the method dispatch
    ```{r}
    setGeneric("DaysInHospital", 
        function(object) standardGeneric("DaysInHospital"))
    setMethod("DaysInHospital", 
        "In-patient", 
        function(object) Sys.Date() - object@DOA)
    DaysInHospital(p3)
    ```

- `callNextMethod()` calls the method for the next class (like `NextMethod()` for S3)

---

class: inverse

## Your turn

Create a new subclass `ICU-patient` which should be a subclass of `In-patient`. This subclass should include a further slot `reason` for the reason of admittance.


- Run `show(p3)` and what you get

- Implement the `show()` method for `In-patient` class, which should print out the DOA in addition to other information. Then redo the last step

- Create an `ICU-patient` object `p5` using `NewPatient3()`. Show the S4 class of the object using `is()`. Make sure to the class is `c("ICU-patient", "In-patient", "Patient")`, and also make sure the object has `pname`, `DOA`, and `reason` fields. 

```{r, include=FALSE, eval=FALSE}
p5 <- NewPatient2(555, 
    "Jack", 
    reason="stroke", 
    DOA = "1965-10-30",
    subclass = c("ICU-patient", "in-patient"))
print(p5)
`print.ICU-patient` <- function(x) {
    NextMethod()
    writeLines(sprintf("Admitted because of %s", attr(x, "reason")))
    invisible(x)
}
print(p5)
```

---

## Directional data example

Here we will implement an S4 class for directional data, namely data that lie on a unit circle

```{r, echo=FALSE, fig.height=5, fig.width=5, fig.align='center'}
set.seed(1)
theta <- runif(10)
x <- rbind(cos(theta * 2 * pi), sin(theta * 2 * pi))
thetaDen <- seq(0, 1, length.out=200)
xDen <- rbind(cos(thetaDen * 2 * pi), sin(thetaDen * 2 * pi))
plot(x[1, ], x[2, ], axes = FALSE, ann=FALSE, xlim=c(-1.1, 1.1), ylim=c(-1.1, 1.1), asp=1)
arrows(0, 0, x[1, ] * 0.9, x[2, ] * 0.9)
lines(xDen[1, ], xDen[2, ])
```

---

- Represent a point by $\theta \in [0, 1)$. E.g., $\theta=1/4$ means 1/4 cycle clockwise (the 3 o'clock direction)
```{r}
setClass("Direction", 
         slots = c(theta = "numeric")) # theta in [0, 1)

# better write a constructor and make sure theta is always a double, not an integer
a <- new("Direction", theta=0) 
```

---

```{r}
setValidity("Direction", function(object) {
  if (all(object@theta >= 0 & object@theta < 1)) {
    TRUE
  } else {
    "theta must lie in [0, 1)"
  }
})
eps <- sqrt(.Machine$double.eps)
d1 <- new("Direction", theta=seq(0, 1 - eps, length.out=4))
d1
```
---

- Create a new generator
    ```{r}
    RUnifDirection <- function(n) {
      new("Direction", theta = runif(n, max = 1 - eps))
    }
    d2 <- RUnifDirection(3)
    ```

---

## Implementing the arithmatics

We will be doing some arithmatics on the directions. `x + 0.3` should mean rotating `x` by 0.3 cycles clockwise

- S4 method can dispatch on any number of arguments. This is called *multiple dispatch*
    ```{r}
    AddNumToDirection <- function(num, d1) {
      # follows the recycling rule
      d1@theta <- (d1@theta + num) %% 1
      d1
    }
    setMethod(`+`, c("Direction", "numeric"), 
              function(e1, e2) AddNumToDirection(e2, e1))
    setMethod(`+`, c("numeric", "Direction"), 
              function(e1, e2) AddNumToDirection(e1, e2))
    setMethod(`+`, c("Direction", "Direction"), 
              function(e1, e2) AddNumToDirection(e1@theta, e2))
    ```
---

```{r}
d1 + 0.3
d1 + d1
```

---

## Inheritance

- We want to a new superclass representing a direction field. Namely, it implements directions starting from arbitrary locations (instead of just the origin)

- Firstly, define a new class which is a location in 2D
    ```{r}
    setClass("Location", 
             slots = c(loc = "matrix"),
             prototype = list(loc=matrix(NA_real_, nrow=2, ncol=0)))
    l1 <- new("Location")
    l2 <- new("Location", loc= matrix(as.numeric(1:6), 2, 3))
    ```


---

- Then define the superclass, which inherits from two classes. S4 objects can inherit from multiple superclasses. This is called *multiple inheritance* 
    ```{r}
    setClass("DirectionField",
             contains = c("Direction", "Location"))
    dl2 <- new("DirectionField", d2, l2)
    dl2
    ```

---
- We can define the `length()` method for the superclasses
    ```{r}
    setMethod("length", "Direction", function(x) length(x@theta))
    setMethod("length", "Location", function(x) length(x@loc))
    length(l2)
    length(d1)
    ```

- It works for the `DirectionField` subclass, but which method are we using?
    ```{r}
    length(dl2)
    ```

- .style[Avoid this kind of ambiguous specification by providing a specific method]


---

## Getters and setters

- Getters and setters are functions to extract and set information

- They should be provided to the users for the consistency of Application Programming Interface (API)
    - What matters to users is only the interface exposed to the them; the underlying implementation does have need to have anything to do with that
    - The developer may change the underlying implementation without changing the API

---

- Create a getter and a setter for the position of the points in terms of the degree
    ```{r, results='hide'}
    setGeneric("degree", function(object) standardGeneric("degree"))
    setMethod("degree", "Direction", 
              function(object) object@theta * 360)
    setGeneric("degree<-", 
               function(object, value) standardGeneric("degree<-"))
    setMethod("degree<-", "Direction", function(object, value) {
        object@theta <- value / 360
        validObject(object)
        object
    })
    ```

- The setter should prevent the user from setting erroneous values

---

```{r}
degree(d1)
d2 <- d1
degree(d2) <- degree(d2) / 3
d2
```


---

## Example use of S4 in packages

- `stats4::mle`, for the `mle` class and related methods

- `lubridate::period` for date/time calculation

- [`IRanges`](https://bioconductor.org/packages/release/bioc/html/IRanges.html) for integer genomic range manipulation (available on Bioconductor)

---

## Mixing S4 and S3

- You can mix S4 and S3

- Register an S3 class as S4 using `setClass()` and `setOldClass()`

- You can use an S3 generic as an S4 generic. Better yet, call `setGeneric()` to make this explicit

---

class: big, middle

## R6

---

## R6

- S3/S4 are *functional OOP* systems, namely methods belong to generic functions

- R6 is an *encapsulated OOP* system, namely methods belong to a class

- R6 follows reference semantics, so R6 objects are not copied

    - Good for implementing notions that exists only in a single version, e.g., a remote machine, a database connection, etc

- R6 implements public vs private fields & methods for additional reliability

- Load `library(R6)` to use it

---

## An example creating a logger

```{r}
library(R6)
Logger <- R6Class("Logger", 
    public = list(
        filepath = character(),
        log = function(text) {
          writeLines(text, private$con)
          private$nlines <- private$nlines + length(text)
          invisible(self)
        },
        initialize = function(name="log.txt") {
          self$filepath <- name
          private$con <- file(name, open="w")
        },
        finalize = function() {
          close(private$con)
          # file.remove(self$filepath)
        }
    ),
    private = list(
        nlines = 0,
        con = NULL
    )
)

```

---

## Reference sematics

```{r}

myLog <- Logger$new("tmp1.txt")
myLog$log(c("data cleaning...", "training..."))
myLog

```

---

```{r}
f <- function(logger) {
  logger$log("Your model is phenomenal")
  invisible(logger)
}

f(myLog)
myLog
```


---

## Automatic cleanup

- When an R6 object is removed, the finalizer will be called, performing clean-ups (here, closing a file connection)

- `rm()` remove a binding, while `gc()` garbage collects unbound objects

```{r}
rm(myLog)
gc()
```

---

## References

- Chapters 13 and 14, [Advanced R](https://adv-r.hadley.nz/index.html) by Hadley Wickham
