---
title: "S4 and R6"
subtitle: "PH 290"
author: "Xiongtao Dai"
output:
  xaringan::moon_reader:
    css: [myslides.css, myslides-fonts.css]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    nature:
      beforeInit: "macros.js"
      highlightLines: false
      countIncrementalSlides: false
---

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

## Outline

- The S4 system

- The R6 system

---

class: big, middle

## S4

---


---

class: inverse

## Your turn

Create a new subclass `ICU-patient` which should be a subclass of `in-patient`. This subclass should include a further field `reason` for the reason of admittance.

- Create a ICU-patient object `p5` using `NewPatient2()`. Make sure to the class is `c("ICU-patient", "in-patient", "patient")`, and also make sure the object has `pname`, `DOA`, and `reason` fields. 

- Run `print(p5)` and what you get. Use `s3_dispatch(print(p5))` to reason about the the method dispatch

- Implement the `print.ICU-patient()` method, which should print out the reason in addition to other information. Then redo the last step

```{r, include=FALSE, eval=FALSE}
p5 <- NewPatient2(555, 
    "Jack", 
    reason="stroke", 
    DOA = "1965-10-30",
    subclass = c("ICU-patient", "in-patient"))
print(p5)
`print.ICU-patient` <- function(x) {
    NextMethod()
    writeLines(sprintf("Admitted because of %s", attr(x, "reason")))
    invisible(x)
}
print(p5)
```


---

## One more example

Suppose we have proposed a new "Tremendous Regression". The following function returns an `tm` object, which is a subclass of an `lm` model

- Needs a function to construct the class
    ```{r}
    tm <- function(...) {
        res <- lm(...)
        class(res) <- c("tm", class(res))
        res
    }
    fmod1 <- tm(extra ~ group, sleep)
    ```

- Generics like `plot()`, `print()`, `coef()` works right out of the box thanks to S3 method dispatch

---

- Reimplement a more specific method for an existing generic
    ```{r}
    print.tm <- function(x, ...) {
        cat(sprintf("Fitting: %s\n\n", deparse(x$terms)))
        cat("Coefficients:\n")
        print.default(coef(x))
        invisible(x)
    }
    fmod1
    ```

---

- Implement a new generic
    ```{r}
    present <- function(x, ...) {
        UseMethod("present")
    }
    ```

- ... and a method
    ```{r}
    present.tm <- function(x, ...) {
        print(x)
        writeLines(praise::praise())
        invisible(x)
    }
    present(fmod1)
    ```

---

## Example use of S3 in packages

- [`caret`](https://cran.rstudio.com/web/packages/caret/), wildly popular for classification and regression training

- [`RFPCA`](https://github.com/CrossD/RFPCA), for Riemannian and multivariate functional data analysis


---

## References

- Chapters 12 and 13, [Advanced R](https://adv-r.hadley.nz/index.html) by Hadley Wickham
