---
title: 'Homework #2'
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Your submission should include both code and well-formatted explanation (not just R comments) and is expected to be reproducible. You may decide for yourself what reproducible file format do you use. 
*Please post on Ed Discussion (accessible in bCourses) for technical questions.*

## Code review

We will be reviewing other's code for Question 4 in [HW1](hw1.html). Each of you will be the reviewer who reviews another classmate's code, and also the coder who send out the code for review. Please find the reviewer assignment list [here](codeReview.xlsx).

1. (Code review Part I: Coder) 
Provide the following to the assigned Reviewer:

    - Source code
    - Provide at least one example usage of your function
    - Documentation, detailing the usage of your function (Also make sure your function does what it describes)

    Please send your files through email to the assigned reviewer and copy the [instructor](mailto:xdai@berkeley.edu). The email should be titled exactly "PH 290 HW2 Code Review".
    **Please finish this step before this homework is due.**
    

1. (Code review Part II: Reviewer) Write 2-3 pargraphs to comment on correctness, reproducibility, clarity, and coding style of the code you reviewed. Make sure to be kind and constructive to your classmate 😊. In addition to the paragraphs, please include in your report

    - come up with at least three new usages of the function (three input scenarios) to test out the function, possibly using new datasets; and 
    - highlight three strengths and three risks in the code.

    The code review report should be a separate report from your usual HW2 submission. Please send the report by replying all the previous email when you are done. **This part is due 7 days after the due date of this homework.** 

## Scoping rules


1. Reason with each of the following code (run them in your head). Make a prediction of the output before running the code (note that the last two questions may have multiple outputs). Use the principles of R scoping rules to explain why the output makes sense.

    a. 
        ```{R, eval=FALSE} 
        x <- 10
        f1 <- function(x) {
          function() {
            x + 10
          }
        }
        g <- f1(1)
        g()
        ```
    
    b. 
        ```{r, eval=FALSE}
        y <- 2
        i1<- function() {
          z <- 3
          c(y, z)
        }
        y <- 3
        i1()
        ```    
    
    a. 
        ```{r, eval=FALSE}
        f <- function(x) print("ha")
        g <- function(y) print("hello")
        y = f()
        g(y)
        ```
        
    a. 
        ```{r, eval=FALSE}
        f <- function(x) print("ha")
        g <- function(y) print("hello")
        g(y = f())
        ```



1. We are going to *numerically* calculate the definite integral 
$$\int_l^u f(x) dx$$
of an arbitrary polynomial
$$f(x) = a_0 + a_1 x + a_2 x^2 + \dots + a_n x^n, \quad x\in \mathbb{R}$$
for $n \ge 0$. 
The input is a vector `a` specifying the coefficients (`a[1]`=$a_0$, `a[2]`=$a_1$, etc), a lower bound `lower` ($l$), and an upper bound `upper` ($u$). 

    Follow the following steps:

    a. create a function factory, namely, a function that takes in the coefficients `a` and outputs a polynomial function. E.g., if you input `a = c(1, -1)`, the returned value should be a function that does the same as `function(x) -1 + 1 * x`. Make sure the manufactured function is vectorized, namely if you input a vector to a manufactured function, the output is a vector with the same length
    a. Test that your function factory behaves as expected. Write at least one unit test for each of the following aspects:
        i. The output of your function factory is a function
        i. Manufacture a function with $a_0 = 2$. Given a vector input `x`, the manufactured function returns a vector of all `2` with the same length as `x`
        i. Manufacture a function with $a_0 = 1$ and $a_1 = -1$. The manufactured function works well given a vector input `x`
    a. Manufacture a polynomial function with $a_0 = 1$, $a_1 = 0$, and $a_2 = 1$, and apply the `integrate` function to evaluate the definite integral of the manufactured function from 0 to 10 (e.g., `integrate(f, 0, 10)`)
